Modeling tabular icebergs \textcolor{red}{submerged in} \sout{coupled to} an ocean model

### 2017-06-13 submission1/Submission2_prep/Response_to_reviews.tex

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amscd}
\usepackage{fancyheadings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{float}
\usepackage[margin=2.5cm]{geometry}

\usepackage{color}

\definecolor{light-gray}{gray}{0.75}
\definecolor{dark-gray}{gray}{0.35}

\floatstyle{boxed}
\restylefloat{figure}
\color{black}
\usepackage{lineno}
\color{black}

%\linenumbers

\begin{document}
Dear Editor,\\
\\
\\
Below are point-by-point responses covering the reviewers' individual comments. A copy of the manuscript has also been included with changes and additions recorded in red. \\
\\
Sincerely,\\
Alon Stern, on behalf of the co-authors.
\clearpage
\textbf{Reviewer 1 Evaluations:}\\
Recommendation: Return to author for minor revisions\\
Significant (Required): Yes, the science is at the forefront of the discipline.\\
Supported (Required): Yes\\
Referencing (Required): Yes\\
Quality (Required): The organization of the manuscript and presentation of the data and results need some improvement.\\
Data (Required): No\\
Accurate Key Points (Required): Yes\\
\\
Reviewer 1 (Formal Review for Authors (shown to authors)):\\
\\
The article presents a novel and very exciting new methodology to include tabular icebergs into general circulation models. The new approach uses numerical bonds to group single icebergs/ice elements with finite extent together to form tabular icebergs with complex shapes, where previously levitating point-particle icebergs with typically zero area coverage have been used. The article is an original and exciting contribution in the ongoing discussion about the inclusion of giant tabular icebergs into ocean and climate models, and therefore -in my opinion- very suitable for publication inÊJAMES.Ê\\
\\
\textcolor{red}{}\\
\\
The authors describe the model and the technical details very well, and there are only some points that should be worked out in more detail (see specific comments below). This concerns (a) details about the surface areas where the melt rates and various forces are acting on when proceeding from rectangular point-particle icebergs to (bonded) hexagonal ice elements. In contrast to the thermodynamics/melting parameterization, it is also less clear how (b) the dynamics of inner ice elements is handled, and whether these could be subject to erroneous form drags despite being actually "shielded" by means of the outer elements. Finally, (c) when exactly do the interactive forces enter the computation? Are the ice elements first allowed to drift independently, and then the velocity/position is corrected for via $F_{ia}$ (or $a_{ia}$) afterwards? This could all be suitably added to the appendix; it is important information for other model centers that apply point-particle iceberg models and which could desire to also extend their existing models with interactive bonding forces.\\
\\
\textcolor{red}{}\\
\\
As a comment, despite some obvious issues when using point-particle methods for the modeling of large tabular icebergs, there have been previous successful attempts to model tabular icebergs with the existing technology (after some modifications), and this could be mentioned in your review-like introduction as well (see specific comment below). Those studies are, i.a., Lichey and Hellmer (2001), Hunke and Comeau (2011), and Rackow et al. (2017), which are already cited in the manuscript except for Hunke and Comeau (2011).Ê
The point-particle method has been modified in the above studies in order to improve the drift representation of tabular icebergs, e.g., by (i) including sea ice-iceberg interaction, leading to increased drift speeds when icebergs are "captured" in sea ice, and (ii) by modifications to the sea surface slope force $F_{ss}$ in order to account for the vast horizontal extent of tabular icebergs, and (iii) by manually embedding icebergs into the ocean vertically, as was also done in Merino et al. (2016) (cited in l. 255).
\\
\textcolor{red}{A paragraph has been added to the introduction at line 81 which reviews some of the attempts to move beyond levitating point particle icebergs using the existing technology.} \textcolor{blue}{[Lichey and Hellmer (2001) is not discussed in details since the suggested parametrization is not related directly to the size and structure of tabular icebergs. It is cited as a paper that has modeled tabular icebergs using existing technology.]}\\
\\
Your proposed model framework is a very original idea and is able to produce impressive iceberg-iceberg and iceberg-coastline interactions, but the added value of resolving realistic collisions and ocean-embedded icebergs in a climate context is not immediately clear, and so it is a little too early to entirely deny the applicability of the simpler point-particle model to the tabular iceberg case. At some point, it would be great to see a comparison of the new methodology with the "old" one and to show the possible superiority of the new approach in a systematic manner.\\
\\
\textcolor{red}{}\\
\\
Overall, the study presents fabulous work and, I think, the fact that icebergs are finally non-levitating should be strongly emphasized, also in the abstract, because it is a real milestone as you rightly mention in lines 467-469. I recommend the paper for publication in JAMES subject to minor revisions.\\
\\
\textcolor{red}{We would like to thank the reviewer for the very favorable review of our work. We agree that the fact that icebergs are non-levitating in our new iceberg framework is an important feature which was underemphasized. We have added the word `submerged' to the paper title, and edited the abstract, introduction and conclusion to further emphasize this point.\\
\\
Line 18 in abstract has been edited and now reads: ``In this framework, tabular icebergs are represented by non-levitating Lagrangian elements that drift are submerged in the ocean.'' \\
\\
A brief discussion of levitating and non-levitating icebergs models has been added to the new paragraph in the introduction at line 81. The development of a iceberg model where icebergs are submerged in the flow (and not levitating) has been added as one of the goals of the study at line 103.\\
\\
A few words have been added to the first paragraph of the conclusion to emphasize that the iceberg models allows icebergs to be non-levitating.}\\
\\
----
Some general clarifications would further improve the paper:\\
\\
\textcolor{red}{}\\
\\
- When initializing your idealized ice shelf, the ice elements will have decreasing thickness towards the ice shelf front, correct? If the ice elements behave like icebergs in point-particle models, that would also imply different freeboards. Do you somehow enforce a level ice shelf surface? Do you generally enforce a level surface for bonded ice elements, i.e. tabular icebergs?\\
\\
\textcolor{red}{All of the ice elements are assumed to be floating in the ocean (in isostatic equilibrium), so that the mass of a column of ice is equal to the mass of water it displaces.  This means that ``the element thickness is related to the draft and freeboard by $T=F+D$ and $D=\frac{\rho}{\rho_{o}} T$'' (see line 610). Since the ice shelf is thickest at the grounding line, and gets much thinner toward the ice front, this implies that the the elevation of the ice surface decrease towards the ice front (as the reviewer correctly points out).\\
\\
The following lines have been added at line 410 to clarify: ``The initial ice thicknesses of the elements are calculated from the ice mass by assuming that the ice has a constant density, $\rho$. The draft and freeboard are related to the ice thickness by $T=F+D$ and $D=\frac{\rho}{\rho_{o}} T$, where $\rho$ and $\rho_{o}$ are the density of ice and seawater respectively."} \textcolor{blue}{[Perhaps we should not add these lines to the text since they already appear in the text in the appendix? ]}\\
\\
- In the case of tabular icebergs, your bonded ice elements are ice columns that are potentially close to the limit for iceberg capsizing. Do you turn off the capsizing parameterization for bonded ice elements? Would the ice elements capsize immediately after bond breakup?\\
\\
\textcolor{red}{1) Iceberg rolling is turned off bonded icebergs. 2) All icebergs are initialized with D>??? and W=?. This means that W/D > 0.92 stability requirement. (hopefully).}\\
\\
- Have you already followed a tabular iceberg until it was completely melted? What about capsizing of a laterally strongly eroded former tabular iceberg that still consists of several bonded ice elements, is this even possible?Ê\\
\\
\textcolor{red}{}\\
\\
- From your experience, will the timestep limitation due to the interactive forces be too prohibitive for long climate model runs? What is the computational overhead when using the new framework compared to the original point-particle model code?\\
\\

----
Specific comments:\\
\\
l. 14: comma after "sea ice formation"\\
\\
\textcolor{red}{Comma has been added}\\
\\
l. 16: "the current generation of ocean circulation models do not represent" -> "the current generation of ocean circulation models usually does not represent"\\
\\
\textcolor{red}{The word `usually' has been added.}\\
\\
l. 22: I think the comma after "its breakup" should be dropped\\
\\
\textcolor{red}{Comma has been removed}\\
\\
l. 60-64: It is fair to say that these models have been mostly used for icebergs smaller than 3.5 km on a global scale, but there are also studies applying the point-particle models to tabular Antarctic icebergs (see above comment). I suggest to extend your short overview by also discussing those simpler models in order to put the new framework into a broader context.\\
\\
\textcolor{red}{A paragraph has been added to the introduction at line 81 which reviews some of the attempts to move model tabular icebergs using the existing technology.} \\
\\
l. 72-74: As far as I understand, iceberg breakup and calving is still an unresolved issue even when using bonded ice elements? The technical capability to cut an iceberg into pieces is certainly there, but the "when" and "how" is still unresolved (which bonds should be removed and why/when?). One could argue that dynamic iceberg lists in simple point-particle models already provide a similar technical capability, because an iceberg that is destined to break into pieces could simply be removed from that list and be replaced by a set of smaller child icebergs. I have the impression that not much is gained using the new methodology with regard to the exact timing/modeling of iceberg calving and break up. Maybe you can clarify the sentence and what you mean by "representation of iceberg breakup and calving".\\
\\
\textcolor{red}{We agree with the reviewer that the issue of iceberg breakup and calving is still unresolved and there is still much to be understood about the physic that govern these processes.  We do however believe the the present work provides a framework which allows iceberg calving (away from a submerged ice shelf) to be modeled more easily. Firstly, is not clear how the calving of icebergs away from a submerged ice shelf could be represented using levitating icebergs (since removing mass from the ice shelf would trigger artificial tsunamis). If icebergs are non-levitating then it becomes important that the size of the iceberg matches the size of the piece of ice calving away from the ice shelf. 
Calving icebergs can have many different structures and shapes. It is not clear how this could be represented using rectangular, circular icebergs or point particle icebergs.
Representing icebergs and ice shelves using ice elements submerged in the ocean means therefore puts fewer restrictions on possible calving laws. Similarly with icebergs breakup, it is unclear how one would represent an iceberg breaking up into two pieces with distinct shapes by introducing child icebergs to replace the parent. This could be achieve by keeping track of icebergs shape, but this is a non-trivial exercise.\\

Lines 72-74 have been removed, and have been replaced with a very brief discussing of how levitating point particle icebergs are largely incompatible
with submerged ice shelf models.}\\
\\
l. 84: "should run sufficiently quickly" $->$ "should run sufficiently quick" (or fast)ÊÊ\\
\\
\textcolor{red}{Changed to "should run sufficiently fast"}\\
\\
l. 139-143: What happens to bonded ice elements? The body forces $F_{ss}$ and $F_{c}$ acting on an ice element are clearly not affected by the number of neighboring ice elements, but I suspect you would treat the other forces acting on the inner ice elements differently, probably similar to the modifications for the melting paramerizations in Section 2.5 (eq. 9 and 10)?\\
This is because inner ice elements should be subject to atmospheric and oceanic skin (surface and bottom) drags only, and the form drags would only act on outer ice elements; ice elements with less than 6 neighbors could partially feel form and skin drags. Similarly, wave radiation and sea ice drag should only act on the outer elements, correct? It wasn't clear to me how this is handled, also after reading appendix A. Moreover, which areas are the forces acting on in extension to the Martin and Adcroft (2010) case? Are they formulated based on the apothem $A_{p}$?\\
\\
\textcolor{red}{}\\
\\
l. 174: Later, you are mostly using "$d_{i,j}$" with comma instead of "$d_{ij}$". Similarly, $L_{ij}$ in l. 178 should be changed consistently.\\
\\
\textcolor{red}{The comma between i and j has been removed throughout the text for consistency.}\\
\\
l. 186: The force formulation in the first two rows of equation 6 is identical, so what about combining it into a single case using if (...) OR if (... and ...) ?\\
\\
\textcolor{red}{}\\
\\
l. 191-194: I was worried that bonds between ice elements of very different size and mass could result in strange behaviour, if you choose your timestep too large, and strong oscillations around the center of mass. After some thought, I think this is a very clever way to deal with the fact that tabular icebergs (bonded ice elements) will not melt homogeneously, but instead will erode very quickly along the outer sides, leading to high accelerations of the (smaller and lighter) ice elements along the iceberg boundary compared to the slower melting inner ice elements. If noteworthy, maybe you can elaborate a bit more about your experiences leading to that adjustment.\\
\\
\textcolor{red}{}\\
\\
l. 200: add vector arrow to $r_{ij}$\\
\\
\textcolor{red}{}\\
\\
l. 202: "is used reduce", change to "is used to reduce"\\
\\
l. 213: Is it possible to briefly derive the stability condition?Ê\\
\\
\textcolor{red}{}\\
\\
l. 221-225: Very impressive! Why does it rotate exactly? Is it because the first element hitting the coast is stopped (how is this handled?), and due to the bonds the other elements are forced onto a circular path? By the way, are the coastal points also some kind of "stationary ice elements" in your simulation, or how would this be handled in global simulations? Would you need to prescribe repulsive forces along the coasts?\\
\\
\textcolor{red}{}\\
\\
l. 236: "is the length the apothems" -> "is the length of the apothems"\\
\\
\textcolor{red}{}\\
\\
l. 246, 248, 387, 407, 439, ... : Please change "Eularian" to "Eulerian" throughout the text, there are many occasions where it is misspelled.\\
\\
\textcolor{red}{}\\
\\
l. 250: delete ":"\\
\\
\textcolor{red}{}\\
\\
l. 254/255: Similarly to the vertical embedding, there is no need to manually embed tabular icebergs -horizontally- into the ocean by averaging $F_{ss}$ over a larger area, as suggested in above comment (ii), because in your framework the sea surface slope force $F_{ss}$ is conveniently evaluated in different locations (for every ice element separately).\\
\\
\textcolor{red}{}\\
\\
l. 260: "(iii) imposing heat, salt, and mass fluxes on the ocean, associated with ice melting" Which salinity do you assume for the iceberg meltwater?\\
\\
\textcolor{red}{}\\
\\
l. 277: "[...] This method allows for the intersection to be found even when the hexagon is not aligned with the grid." Is your implementation particularly efficient so that it could be worth sharing it here?\\
\\
\textcolor{red}{}\\
\\
l. 282: comma after "i.e." instead of ":"\\
\\
\textcolor{red}{}\\
\\
l. 287: "described" -> "describe"\\
\\
\textcolor{red}{}\\
\\
l. 295: "The details of $M_{b}$, $M_{v}$, and $M_{e}$ are given in Appendix A." What is "W" for an ice element? Do you assume hexagons or circles? Is it 2*$A_{p}$?
Moreover, what is the area that, for example, the wave erosion $M_{e}$ is acting on (for rectangular point-particle icebergs this is usually assumed to be 2 side walls, I think; what is it for hexagons, 3 side walls?) It might prove to be very useful to list the surface areas the different forces and melt rates are acting on in the new framework, where you proceeded from point-particle icebergs to (hexagonal) ice elements.\\
You could list it either here or in the appendix.\\
\\
\textcolor{red}{}\\
\\
l. 302 and 306: delete the comma before "$M_{b}$" and "$M_{s}$"\\
\\
\textcolor{red}{}\\
\\
l. 308: "... which is a typical melting parameterization used beneath ice shelves [Holland and Jenkins, 1999]." One could mention that the 3-equation model for ice shelf melting has been previously applied for the estimation of tabular iceberg melting, e.g. in Silva et al. [2006] and Rackow et al. [2017]. When proceeding to larger iceberg structures this is a natural extension of the model's original range of application.\\
\\
\textcolor{red}{}\\
\\
l. 309-318: This is a clever way to account for the different melting along the boundary elements and inner elements (eq. 9 and 10). Since you do the weighting for the melt rates, the bottom and side areas where the bottom and side melt rates are acting on is unchanged compared to MA2010, correct?\\
\\
\textcolor{red}{}\\
\\
l. 347: "in a following way" -> "in the following way"\\
\\
\textcolor{red}{}\\
\\
l. 348/349: "up to six bonds per element" Do you support more bonds, for example along the ice shelf front where you added smaller elements to fill the gaps (l.403/404)?\\
\\
\textcolor{red}{}\\
\\
l. 360/361: "from one processor the next ..." -> "from one processor to the next, added to and removed from the appropriate lists, and the"\\
\\
\textcolor{red}{}\\
\\
l. 363: "a one" -> either "a" or "one"\\
\\
\textcolor{red}{}\\
\\
l. 389: "in an idealized setting"\\
\\
\textcolor{red}{}\\
\\
l. 400, l. 434, l. 441: delete ":" in section heading\\
\\
\textcolor{red}{}\\
\\
l. 405: "preprocessing inversion" What is that supposed to mean?\\
\\
\textcolor{red}{}\\
\\
l. 418: "using the ALE regridding-remapping scheme"\\
\\
\textcolor{red}{}\\
\\
l. 468: "modeling" -> "model"\\
\\
\textcolor{red}{}\\
\\
l. 495: "tabular icebergs"\\
\\
\textcolor{red}{}\\
\\
l. 515: "As the iceberg drifts"\\
\\
\textcolor{red}{}\\
\\
l. 532: "phenomenon"\\
\\
\textcolor{red}{}\\
\\
l. 540/541: Surprisingly, this is in contrast to the findings in Silva et al. [2006], where the authors show that the point-particle iceberg-melt parameterization leads to "on average, half the amount of melting compared to the Holland and Jenkins [1999] model." What could be the reason for this?\\
\\
l. 570: "at the edge"\\
\\
\textcolor{red}{}\\
\\
l. 572: "showed" -> "show"\\
\\
\textcolor{red}{}\\
\\
l. 574: "which is an important process"\\
\\
\textcolor{red}{}\\
\\
l. 580: "the question of how and when to introduce tabular icebergs into the ocean needs to be addressed." You could nicely refer to your earlier Stern et al. [2016] paper, where this question is explicitly raised. The question was also echoed and discussed in the Rackow et al. study.\\
\\
\textcolor{red}{}\\
\\
----
Comments on the appendix:\\
\\
l. 602: "the velocities of air,"\\
\\
\textcolor{red}{Corrected}\\
\\
l. 604 "freeboard,"\\
\\
\textcolor{red}{Corrected}\\
\\
l. 621: "as a result of wave erosion" (delete melt due to)\\
\\
\textcolor{red}{Corrected}\\
\\
l. 623 "wave erosion 'melt' rate", l. 631 "the 'melt' due to wave erosion"\\
\\

### 2017-06-13 JAMES_iceberg_only.tex
Working on JAMES revisions and Lagrangian paper

\title{Modeling tabular icebergs coupled to an ocean model}

@@

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering ambient ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models do not represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are represented by Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
%We describe the Lagrangian bonded iceberg model and discuss a number of the technical elements involved in its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg drifting away an idealized ice shelf.} 
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf) and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and its breakup, demonstrate capabilities of the developed framework.
\end{abstract}

@@

In recent years, there has been an increased interest in iceberg drift and decay. This surge of interest has been driven by (i) the need to understand polar freshwater cycles in order to create realistic climate change and sea level projections \citep{Silva2006, Shepherd2007, Rignot2013}; and (ii) the increased navigation and exploration activities in high-latitude iceberg-filled waters in the Arctic \citep{Pizzolato2014, Unger2014, Henderson2016}. The increased interest in icebergs has led to the development of numerical models of iceberg drift and decay \citep{Mountain1980, Bigg1997, Gladstone2001, Kubat2005}, some of which have been included in global General Circulation Models \citep{Martin2010, Marsh2015}. These iceberg drift models treat icebergs as Lagrangian point particles, which are advected by the flow, and melt according to parameterizations for icebergs melt. Since icebergs are treated as point particles, iceberg drift models are mostly suitable for modeling icebergs smaller than an ocean grid cell. Consequently, these models have mostly been used to represent icebergs smaller than 3.5~km on a global scale \citep{Jongma2009, Martin2010, Marsh2015}.

Point-particle iceberg drift models are less suitable for modeling larger tabular icebergs, where the size and structure of the iceberg may be an important feature in determining their drift and decay \citep{Stern2016}. They also are not suitable for studying the local effects that icebergs have on the surrounding ocean, or the small scale processes that influence iceberg melt and decay \citep{Wagner2014, Stern2015}. For these reasons, tabular icebergs are currently not represented in the iceberg drift models used as components of climate models, despite accounting for the vast majority of the total Southern Hemisphere iceberg mass \citep{Tournadre2016}. Point-particle iceberg models also do not have any representation of iceberg breakup and calving, which is known to be an important iceberg decay mechanism that influences iceberg trajectories.
%Similarly, operational iceberg models are often inadequate for projecting iceberg drift at fine scales \citep{Turnbull2015}.

The goal of this study is to develop a new framework to model all kinds of icebergs, where tabular icebergs are explicitly resolved in the ocean. Our new representation of icebergs aims to include the following key properties: (i) icebergs should be able to travel large distances within the ocean, (ii)  icebergs should melt and decay as they drift in the ocean, (iii) icebergs should behave as if they have finite extent (in order to study local effects that icebergs have on the surrounding ocean), and (iv) tabular icebergs should be able to break away from ice shelves or break into smaller pieces. Properties (i) and (ii) are common to point-particle icebergs models, while properties (iii) and (iv) are new to the framework developed in this study. A further requirement of the new framework is that the model should run sufficiently quickly to be used in general circulation models used for climate.

@@

To describe the forces between two elements, we begin by introducing some notation.  Let $\vec{x_{i}}$, $\vec{x_{j}}$ be the positions of elements $i$ and $j$. The distance between elements $i$ and $j$ is
\begin{equation}
\label{distance}
d_{ij}= | \vec{x_{i}} - \vec{x_{j}}|.
\end{equation}
In calculations of the interactive forces between elements, the elements are assumed to be circular. We define the interaction radius of an element by 
\begin{equation}
\label{Diameter}
R_{i}=\sqrt{\frac{A_{i}}{\pi}},
\end{equation}
where $A_{i}$ is the planar surface area of element $i$. Using this, we define the critical-interactive-length scale, 
\begin{equation}
L_{ij}=R_{i} + R_{j},
\label{eq:Interaction_crit}
\end{equation}
which governs interactions between elements $i$ and $j$. 
Repulsive forces are only applied when $d_{i,j} \textless L_{i,j}$, while for $d_{i,j} \textgreater L_{i,j}$ attractive bonded forces are applied when a bond exists between element $i$ and $j$ (see diagram in Figure \ref{fig:Force_diagram}). 
The interactive forces are designed such that (in the absence of other external forces) bonded particles will settle in an equilibrium position where elements are separated by $L_{i,j}$.

% The bonded forces are designed such that in the absence of other forces, bonded particles will settle in an equilibrium position where elements are separated by the critical interaction length scale $L_{i,j}$.
% (see schematic in Figure \ref{fig:Interactive_forces}).

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements $i$ and $j$ are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F}_{IA})_{ij}$ on an element $i$ by an element $j$ is given by
\begin{equation}
\label{F_interactione}
(\vec{F}_{IA})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements $i$ and $j$.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{i,j}-L_{i,j} \bigg) M_{i,j} \vec{r}_{ij},
\end{equation}
where $\vec{r}_{ij}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element $i$ and $j$, $\kappa_{e}$ is the spring constant, and $M_{i,j}$ is the minimum of the masses of elements $i$ and $j$. The interactive forces obey Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum mass, $M_{i,j}$, is preferred to the average mass, since this means that for two bonded elements a fixed distance apart, the acceleration due to elastic forces is bounded, even when the mass of one of the elements approaches zero.

>>>>

\title{Modeling tabular icebergs \textcolor{red}{submerged in} \sout{coupled to} an ocean model}
%\title{Modeling tabular icebergs submerged in the ocean}

@@

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering ambient ocean circulation, bottom-water formation, sea-ice formation\textcolor{red}{,} and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models \textcolor{red}{usually} do not represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are represented by  \textcolor{red}{non-levitating} Lagrangian elements that \sout{drift}  \textcolor{red}{are submerged} in the ocean\sout{, and} \textcolor{red}{. The elements} are held together and interact with each other via bonds.
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf) and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and its breakup\sout{,} demonstrate capabilities of the developed framework.
\end{abstract}

@@

In recent years, there has been an increased interest in iceberg drift and decay. This surge of interest has been driven by (i) the need to understand polar freshwater cycles in order to create realistic climate change and sea level projections \citep{Silva2006, Shepherd2007, Rignot2013}; and (ii) the increased navigation and exploration activities in high-latitude iceberg-filled waters in the Arctic \citep{Pizzolato2014, Unger2014, Henderson2016}. The increased interest in icebergs has led to the development of numerical models of iceberg drift and decay \citep{Mountain1980, Bigg1997, Gladstone2001, Kubat2005}, some of which have been included in global General Circulation Models \citep{Martin2010, Marsh2015}. These iceberg drift models treat icebergs as \textcolor{red}{levitating} Lagrangian point particles, which are advected by the flow, and melt according to parameterizations for icebergs melt. Since icebergs are treated as point particles, iceberg drift models are mostly suitable for modeling icebergs smaller than an ocean grid cell. Consequently, these models have mostly been used to represent icebergs smaller than 3.5~km on a global scale \citep{Jongma2009, Martin2010, Marsh2015}.

Point-particle iceberg drift models are less suitable for modeling larger tabular icebergs, where the size and structure of the iceberg may be an important feature in determining their drift and decay \citep{Stern2016}. They also are not suitable for studying the local effects that icebergs have on the surrounding ocean, or the small scale processes that influence iceberg melt and decay \citep{Wagner2014, Stern2015}. 
For these reasons, tabular icebergs are currently not represented in the iceberg drift models used as components of climate models, despite accounting for the vast majority of the total Southern Hemisphere iceberg mass \citep{Tournadre2016}. 
\textcolor{red}{Levitating point-particle iceberg are also largely incompatible with current models or ice shelves and ice shelf cavities. Since ice shelf models typical allow the ice shelf to displace the water in the water column, it is unclear how one would calve a piece of ice  (with finite extent and mass) away from an ice shelf to become a levitating icebergs without triggering artificial tsunamis that would cause the model to crash.}
\sout{Point-particle iceberg models also do not have any representation of iceberg breakup and calving, which is known to be an important iceberg decay mechanism that influences iceberg trajectories.}
%Similarly, operational iceberg models are often inadequate for projecting iceberg drift at fine scales \citep{Turnbull2015}.

\textcolor{red}{Some efforts have been made to represent tabular icebergs using levitating point particles and to account for the physical presence of the tabular iceberg by introducing additional parametrizations.}  \citep{Lichey2001, Hunke2011, Rackow2017}
\textcolor{red}{A promising approach involves integrating ocean properties over the implied iceberg surface area} \citep{Rackow2017} \textcolor{red}{or implied iceberg depth} \citep{Hunke2011, Merino2016} \textcolor{red}{(when calculating the iceberg forcing) to account for the horizontal and vertical iceberg extent. This approach could be extended to apply iceberg melt fluxes over a surface area and depth consistent with the implied iceberg shape. However, this approach still does not account for influence that a submerged iceberg can have on the surrounding ocean due to its physical presence in the water column.} \citet{Martin2010} \textcolor{red}{attempt to account for the effects of icebergs being submerged in the ocean by allowing their icebergs to apply a pressure to top-most layer of their layered ocean model. This approach effectively allowed the iceberg to displace the ocean, therefore making the point-particle icebergs non-levitating. However, a limitation of this approach is that it the maximum iceberg size was constrained to be smaller than one ocean grid cell, which meant that tabular icebergs could not be represented} \citep{Stern2016}.

The goal of this study is to develop a new framework to model all kinds of icebergs, where tabular icebergs are explicitly resolved in the ocean. Our new representation of icebergs aims to include the following key properties: (i) icebergs should be able to travel large distances within the ocean, (ii)  icebergs should melt and decay as they drift in the ocean, (iii) icebergs should behave as if they have finite extent \textcolor{red}{and have arbitrary shape and size} (in order to study local effects that icebergs have on the surrounding ocean), \textcolor{red}{ (iv) icebergs should be submerged in the ocean so they are not longer levitating}, and (\sout{i}v) tabular icebergs should be able to break away from ice shelves or break into smaller pieces. Properties (i) and (ii) are common to point-particle icebergs models, while properties (iii) \textcolor{red}{iv and v} \sout{and (iv)} are new to the framework developed in this study. A further requirement of the new framework is that the model should run sufficiently \sout{quickly} \textcolor{red}{fast} to be used in general circulation models used for climate.

@@

To describe the forces between two elements, we begin by introducing some notation.  Let $\vec{x_{i}}$, $\vec{x_{j}}$ be the positions of elements $i$ and $j$. The distance between elements $i$ and $j$ is
\begin{equation}
\label{distance}
d_{ij}= | \vec{x_{i}} - \vec{x_{j}}|.
\end{equation}
In calculations of the interactive forces between elements, the elements are assumed to be circular. We define the interaction radius of an element by 
\begin{equation}
\label{Diameter}
R_{i}=\sqrt{\frac{A_{i}}{\pi}},
\end{equation}
where $A_{i}$ is the planar surface area of element $i$. Using this, we define the critical-interactive-length scale, 
\begin{equation}
L_{ij}=R_{i} + R_{j},
\label{eq:Interaction_crit}
\end{equation}
which governs interactions between elements $i$ and $j$. 
Repulsive forces are only applied when $d_{ij} \textless L_{ij}$, while for $d_{ij} \textgreater L_{ij}$ attractive bonded forces are applied when a bond exists between element $i$ and $j$ (see diagram in Figure \ref{fig:Force_diagram}). 
The interactive forces are designed such that (in the absence of other external forces) bonded particles will settle in an equilibrium position where elements are separated by $L_{ij}$.

% The bonded forces are designed such that in the absence of other forces, bonded particles will settle in an equilibrium position where elements are separated by the critical interaction length scale $L_{ij}$.
% (see schematic in Figure \ref{fig:Interactive_forces}).

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements $i$ and $j$ are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F}_{IA})_{ij}$ on an element $i$ by an element $j$ is given by
\begin{equation}
\label{F_interactione}
(\vec{F}_{IA})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements $i$ and $j$.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{ij}-L_{ij} \bigg) M_{ij} \vec{r}_{ij},
\end{equation}
where $\vec{r}_{ij}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element $i$ and $j$, $\kappa_{e}$ is the spring constant, and $M_{ij}$ is the minimum of the masses of elements $i$ and $j$. The interactive forces obey Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum mass, $M_{ij}$, is preferred to the average mass, since this means that for two bonded elements a fixed distance apart, the acceleration due to elastic forces is bounded, even when the mass of one of the elements approaches zero.
